On the dynamic assessment of the Local-Maximum Entropy Material Point Method through an Explicit Predictor-Corrector Scheme

### 2020-05-28 _region_.tex
Cambios MP 28-5-2020

\section{Introduction}
\label{intro}
Since the proposal of \acrshort{mpm} by Sulsky {\it  et al.}
(1994)~\cite{Sulsky1994} as a generalization to solids of the \acrfull{flip} method~\cite{Brackbill1986}, its popularity
has increased due to its ability to deal with large strain regime
without suffer mesh distortion inaccuracies.\PNA{One of the main fields where this method is strong is the dynamic one since the original time integration scheme proposed was the \acrfull{fe}~\cite{Sulsky1994}, carried out in a explicit manner. However, in this type or problems, the main instabilities of the original \acrshort{mpm} are even marked.}

\PNA{On the one hand, the main source of instability occurs} when
material points cross cell boundaries. This provoked the development
of other interpolation techniques to overcome this limitation such as
the \acrfull{gimp} method Bardenhagen \& Kober (2004)~\cite{Bardenhagen2004}, which has
demonstrated to have a good performance in the finite deformation
regime. However, in the absence of a regular grid, construction of the
weighting functions is only achieved at considerable effort and
computational cost.  Furthermore, as it is a voxel-based
discretization technique, it is prone to suffer voxel domains overlap
or gaps when the material point mesh becomes irregular, which can
introduce severe inaccuracies as noticed Steffen {\it et al.}
(2008)\cite{Steffen2008}. This is similar to the difficulty
encountered by the finite element methods due to element distortion.
A more robust alternative is the \acrfull{ddmp} method proposed by Zhang {\it et al.}
(2011)~\cite{Zhang2011a}. Unfortunately this method shows an
unsatisfactory behaviour when particle/cell ratio
decreases~\cite{DHAKAL2016301}. Therefore \acrshort{ddmp} a large number of particles is needed for convergence, what
makes the method very expensive. In recent years the employment of
spline-lines as shape functions has gain popularity with the introduction of the B-Spline 
\acrshort{mpm} proposed by Roel Tielen {\it et al.} (2017)~\cite{TIELEN2017265},
this technique allows the employment of unstructured set of notes and
particles. More recently, approximants derived from minimization has
been introduced in to the \acrshort{mpm} framework with the \acrfull{ctls}
reconstruction proposed by Wobbes {\it et al.}
(2018)~\cite{E_Wobbes_2018}. Unfortunately, when particles are spread in
a challenging way, the quality of the \acrshort{ctls} approximation decreases locally.

This document adopts the \acrfull{lme}, or Local \textit{Max-Ent} approximates, as a robust substitute\PNA{of the aforementioned} shape functions in \acrshort{mpm}. First introduced by Arroyo \& Ortiz
(2006)~\cite{Arroyo2006}, it belongs to the class of convex 
approximation schemes and provides a seamless transition between
\acrshort{fem} and meshfree interpolations. The
approximation scheme is based on a compromise between minimizing the
width of the shape function support and maximizing the information
entropy of the approximation. The \acrshort{lme} approximation
may be regarded as a regularization, or \textit{thermalization}, of
Delaunay triangulation which effectively resolves the degenerate cases
resulting from the lack of uniqueness or the triangulation. \acrshort{lme} basis functions possess many desirable properties for
meshfree algorithms. First of all, they are entirely defined by the
nodal set and the domain of analysis. They are also non-negative,
satisfy the partition of unity property, and provide an exact
approximation for related functions~\cite{Arroyo2006}. Important contributions on the Maximum-Entropy have been made by Sukumar and coworkers~\cite{Sukumar15} with Cell-based techniques and the ones carried within the \acrfull{otm} method. The latter methodology has been proven to have a good performance under
the dynamic regime by other researchers, being important the contributions of Li {\it et al.} (2012)~\cite{Li2012} and Navas {\it et al.}
(2018)~\cite{Navas:17b,Navas2018a} in the explicit regime and Navas {\it et al.}~\cite{Navas2016,Navas2016b,Navas:17c} and Wriggers and coworkers~\cite{Wriggers18} with implicit schemes. And more
recently, under \acrshort{mpm} framework, the work made by Wobbes {\it et
  al.}(2020)~\cite{Wobbes2020}.\PNA{The proposed research delves into the benefits of
the regularization parameter, \gls{LME-beta}, and the analogy of the different obtained shape functions derive by the tuning of this parameter and the traditional \acrshort{mpm} ones.}

The aforementioned techniques are devoted to mitigate the
``grid crossing'' error. Nevertheless, in the presence of shock waves spurious,
numerical noises appear despite of the employment of these
techniques~\cite{Tran2019e}. These numerical inaccuracies, also known
as wiggles, arise due to inaccuracies in the time discretization technique.
A simple approach to face those spurious noises lies on the addition of nonphysical damping sources to the equilibrium equations. This
approach has been widely employed in this and many other numerical
techniques. To avoid introducing this nonphysical sources, many
researchers has proposed alternative time integration
schemes which reduce the presence of high frequency noises by
filtering them or increasing\PNA{somehow} the accuracy.\MMP{One of the first attempts of it was
the proposal of an implicit time integration scheme by Guilkey \&
Weiss \cite{Guilkey_2003}. More recently Wang {\it et al.}
\cite{Wang_2016} to downsize this spurious noises add a nonvoiscous
damping to the linear momentum balance equation, and later Charlton
{\it et al.} \cite{Charlton_2017} extended this scheme to the
\acrshort{gimp} approach introducing the \acrfull{igimp}
method. However, the local damping introduced by 
\cite{Wang_2016} can totally over-damp the solution in time-dependent
simulations such as in consolidation process. Within the explicit
time integration schemes, Lu {\it et al.}\cite{LU_2018} introduced the time-discontinuous Galerkin method to control the spurious noises
propagation, and later Tran \& Solowski~\cite{Tran2019e}
proposed a generalised-$\alpha$ scheme for \acrshort{mpm} with
promising results but at the expense of increasing the computational
effort.} In this paper a less time consuming and high efficient explicit predictor-corrector integration method has been
proposed. It consists in an accommodation of the traditional \acrsull{npc} scheme, widely employed in Finite Element methods. This method
has been chosen among other suitable alternatives as those proposed
by Wilson {\it et al.} (1972)~\cite{Wilson1972} or Chung \& Hulbert
(1993)~\cite{Geranlized_alpha_1993} because its simplicity and its good
performance dealing with solid dynamic problems under a meshfree
framework in~\cite{Navas2018a}.

@@

To accommodate this step to \acrshort{mpm} framework, it is necessary to get
the nodal values of the velocity and acceleration throughout a variational
recovery process where particles quantities are transferred to the
mesh nodes. This technique arises as a generalization of the super-convergent recovery
procedures described by Zienkiewicz \& Zhu \cite{ZZ1992_I} (\textit{ZZ})
in the context of \acrshort{fem}. In \acrshort{mpm}, Gauss quadratures are not employed. However, 
integrals are computed following the Riemann integral definition,
where each component of the summation corresponds to a particle of the
discretization. Also Bardenhagen \& Kober \cite{Bardenhagen2004}
proved that through this information-transference technique mass and momentum are conserved. So for a general particle variable $\Phi_p$, employing the
\textit{ZZ} technique, it is possible to get its nodal homologous $\Phi_I$  as:
\begin{equation}
  \label{eq:Variational-recovery}
   \Phi_I = \frac{m_p N_{Ip} \Phi_p}{m_I}.
 \end{equation}
 Therefore, to get an analogous expression for
 \eqref{eq:Predictor-velocity-I} in the context of \acrshort{mpm}, the
 procedure described in the equation \eqref{eq:Variational-recovery}
 is employed, obtainen the following expression:
 \begin{equation}
   \label{eq:Predictor-velocity-II}
   \vec{\tilde{v}}_I^{k+1} = \underbrace{\frac{N_{Ip}^{k} m_p
       \vec{v}_p^k}{m_I}}_{\vec{v}_I^{k}} + (1 - \gamma)\ \Delta t\  \underbrace{\frac{N_{Ip}^{k} m_p \vec{a}_p^k}{m_I}}_{\vec{a}_I^{k}}.
 \end{equation}
Nonetheless this way of computing the predictor stage can introduce
instabilities due to numerical cancellation likewise the original
Sulky algorithm. Thankfully, this can be avoided easily by the
equivalent formulation proposed as follows: 
\begin{equation}
  \label{eq:Predictor-velocity-II}
  \vec{\tilde{v}}_I^{k+1} = \frac{ N_{Ip}^{k} m_p (\vec{v}_p^k + (1 - \gamma)\ \Delta t\ \vec{a}_p^k)}{m_I}.
\end{equation}
This way of computing the nodal predictor is both numerically stable
and minimize the computational effort. Once nodal velocities are
obtained, the essential boundary conditions are imposed over \gls{dirichlet-boundary}. And in the
following, the ``classic'' \acrshort{mpm} algorithm continues to reach to the
equilibrium equation \eqref{eq:particle_balance_forces3}. Next, the
\textit{corrector} stage is introduced. Due to the fact that nodal
velocities were obtained earlier, this step is computed in the same way as
in \acrshort{fem},
\begin{equation}
  \label{eq:Corrector-velocity}
  \vec{v}_{I}^{k+1} = \vec{v}_{I}^{pred} + \gamma\ \Delta t\ \frac{\vec{f}_{I}^{k+1}}{\tens{m}_I^{k+1}}.
\end{equation}
Finally updated particle kinetics are computed using nodal values as:
\begin{align}
  \label{eq:Update-lagrangian-pce}
        &\vec{a}_p^{k+1} = \frac{N_{Ip}^k\vec{f}_{I}^{k}}{\tens{m}_I^k}\\
      &\vec{v}_p^{k+1} = \vec{v}_p^n + \Delta t\
        \frac{N_{Ip}^k\
        \vec{f}_{I}^{k}}{\tens{m}_I^k}\\
      &\vec{x}_p^{k+1} = \vec{x}_p^n + \Delta t\
         N_{Ip}^k\ \vec{v}_{I}^{k} +
        \frac{1}{2}\Delta t^2\ \frac{N_{Ip}^k\
        \vec{f}_{I}^{k}}{\tens{m}_I^k}.
\end{align}
Notice that particle displacements are computed using the corrected
nodal velocities as well as the accelerations with the velocities
of the predictor. However, particle velocities and accelerations
are computed using the corrected velocities. Therefore here we share similarities
with the \textit{leapfrog integration} where position is not updated at
full time step, but the velocity is updated at half time steps. Notice
also that, with this approach, the calculation of nodal momentum values
are not required. Due to its simplicity, the proposed scheme can be implemented with minor modifications
over the standard forward Euler. The full implementation is summarized in the algorithm~\ref{algo:1}.

@@

\acrfull{lme} approximation schemes were
first introduced by Arroyo \& Ortiz (2006)\cite{Arroyo2006} and has been
recently tested under \acrshort{mpm} framework by Wobbes {\it et al.}
(2020)\cite{Wobbes2020}.\PNA{The performed simulations of \acrshort{mpm}
within \acrshort{lme} in \cite{Wobbes2020} shows considerably more accurate stress
approximations than traditional \acrshort{mpm} schemes. However, 
how the regularization parameter $\beta$
affects to the accuracy and stability of the solution is not assessed deeply in that research, which is considered interesting in the present manuscript; the tuning of this $\beta$ parameter allows to make the comparison of the accuracy against analogous traditional \acrshort{mpm} shape function.}

@@

application of this capability, in this publication authors smooth 
oscillations due to excess of pore water pressure in consolidation
problems.}A proof of this statements is observed in figure~\ref{fig:LME_MPM}.

@@

This section is devoted to test the ability of both predictor-corrector
time integration scheme and the Local \textit{Max-Ent} approximants to
overcome spurious oscillations due to the grid crossing and high
frequency loads under the context of \acrshort{mpm}. Two different test have
been adopted for this purpose: the benchmark proposed by Dyka \& Ingel
(1995)\cite{Dyka1995} and the test proposed in the PhD thesis of
Andersen (2009)\cite{thesis_Andersen_2009}. On the one hand, the accuracy of the \acrfull{npc}
scheme is compared to the standard \acrfull{fe} in order to assess the performance of the time integration schemes. On the other hand,  \acrshort{lme} solutions
are compared with those provided by \acrfull{ugimp} and Q4 shape
functions. To avoid some mesh-dependent issues, in all
calculations a regular background mesh was set. All simulations were performed with in-hose software.

@@

This benchmark was proposed\PNA{by Dyka~\cite{Dyka1995} since allows so study easily the capability}
of the proposed time integration algorithm to avoid velocity
field instabilities. It consists of a one-dimensional bar of a length of 0.1333 meters, sketched in the
figure~\ref{fig:Dyka_Bar}. The boundary conditions are: displacements are constrained ($\vect{v} \rvert_{x=L} = 0$) in the right
border, letting free in the rest of the body and ($\tens{\sigma} \rvert_{x=0} =
0$). An initial velocity of $\vect{v}_o = - 5\ m/s$ is given to the
left quarter of the bar. Finally, the elastic parameters considered for this test are:
\begin{itemize} 
\item  Density : $7833\ kg/m3$
\item  Poisson ratio : $0$
\item  Elastic modulus : $200 \cdot 10^9\ Pa$
\end{itemize}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}\sidecaption
  \centering
  \resizebox{\hsize}{!}{
    \input{Figures/Dyka_bar.tex}}
  \caption{Geometrical description of the Dyka \cite{Dyka1995} bar.}
  \label{fig:Dyka_Bar}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In this case, 0.0001 seconds\PNA{are considered }in the simulation. Therefore, the elastic wave generated travel through the bar
(from A to C and back to A) at least two times. For the spatial
discretization, a set of seven nodal mesh sizes (0.1, 0.3325, 0.5,
1.0, 3.3325, 6.665, 10.0 millimeters) are considered. For each element a number of
four particles was selected. In the initial layout, particles are
occupying the exact quadrature points of a linear quadrilateral, with
the exception of the \acrshort{ugimp} simulation, where gaps or overlap between
voxels of each particle are not allowed. In those cases, each particle
occupies the center of each cell quarter. For all simulations, time step
is controlled by a Courant-Friedrichs-Levy condition of 0.1, were the adopted
celerity is computed as:
\begin{equation}
  \label{eq:Cel}
  Cel = \max\{\max_{p \in \Omega_p}\{ \vect{v}_p \} , \max_{p \in \Omega_p}\{ \sqrt{\frac{E_p}{\rho_p}} \} \}.
\end{equation}
An important consideration regarding modellization concerns to the
background mesh. Notice that free border of the bar has a maximum
horizontal displacement of 0.03 millimeters, therefore 
a computational domain with an extra gap of 0.03 millimeters is
required in order to accommodate the unconstrained displacement of the
particles in the left border of the bar. Naturally this problem arises
when the mesh size is small enough that relative displacement of the
particles is larger then the distance to the border, so grid crossing
phenomena could appear even in those cases with infinitesimal
displacements.

@@

A first comparative between the velocity results of both time integration
schemes is plotted in figure \ref{fig:Dyka-NPC-FE}. It demonstrates the
superior performance of the \acrshort{npc} \textit{versus} the \acrshort{fe}. In the \acrshort{npc} the
spurious oscillations are quickly mitigated in the first time steps,
and the error is not propagated in time, opposite to the results of the \acrshort{fe} scheme, where the
simulation becomes unstable after $6E^{5}$ seconds.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}\sidecaption
  \centering
  \resizebox{0.9\hsize}{!}{
    \includegraphics[width=\textwidth]{./Figures/Velocity_FE_vs_PCE_CFL_01}
  }
  \caption{Comparative of the \acrshort{npc} \textit{versus} the \acrshort{fe}. In the
    picture the velocity evolution at the point in the bar left side
    is plotted.}
  \label{fig:Dyka-NPC-FE}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

>>>>

\section{Introduction}
\label{intro}
Since the proposal of \acrshort{mpm} by Sulsky {\it  et al.}
(1994)~\cite{Sulsky1994} as a generalization to solids of the \acrfull{flip} method~\cite{Brackbill1986}, its popularity
has increased due to its ability to deal with large strain regime
without suffer mesh distortion inaccuracies.\PNA{One of the main
  fields where this method is strong} \MP{in solid dynamics \PNA{since the original time integration scheme proposed was the \acrfull{fe}~\cite{Sulsky1994}, carried out in a explicit manner. However, in this type or problems, the main instabilities of the original \acrshort{mpm} are even marked.}

\PNA{On the one hand, the main source of instability occurs} when
material points cross cell boundaries. This provoked the development
of other interpolation techniques to overcome this limitation such as
the \acrfull{gimp} method Bardenhagen \& Kober (2004)~\cite{Bardenhagen2004}, which has
demonstrated to have a good performance in the finite deformation
regime. However, in the absence of a regular grid, construction of the
weighting functions is only achieved at considerable effort and
computational cost.  Furthermore, as it is a voxel-based
discretization technique, it is prone to suffer voxel domains overlap
or gaps when the material point mesh becomes irregular, which can
introduce severe inaccuracies as noticed \MP{by} Steffen {\it et al.}
(2008)\cite{Steffen2008}. This is similar to the difficulty
encountered by the finite element methods due to element distortion.
A more robust alternative is the \acrfull{ddmp} method proposed by Zhang {\it et al.}
(2011)~\cite{Zhang2011a}. Unfortunately this method shows an
unsatisfactory behaviour when particle/cell ratio
decreases. Therefore \acrshort{ddmp} a large
number of particles is needed for convergence~\cite{DHAKAL2016301}, what makes the method
very expensive. In recent years the employment of spline-lines as shape functions has gain popularity with the introduction of the B-Spline 
\acrshort{mpm} proposed by Tielen {\it et al.} (2017)~\cite{TIELEN2017265},
this technique allows the employment of unstructured set of notes and
particles. More recently, approximants derived from minimization has
been introduced in to the \acrshort{mpm} framework with the \acrfull{ctls}
reconstruction proposed by Wobbes {\it et al.}
(2018)~\cite{E_Wobbes_2018}. Unfortunately, when particles are spread
in \MP{some particular special patterns}, the quality of the \acrshort{ctls} approximation \MP{decrease} locally.

This document adopts the \acrfull{lme}, or Local \textit{Max-Ent} approximates, as a robust substitute\PNA{of the aforementioned} shape functions in \acrshort{mpm}. First introduced by Arroyo \& Ortiz
(2006)~\cite{Arroyo2006}, it belongs to the class of convex 
approximation schemes and provides a seamless transition between
\acrshort{fem} and meshfree interpolations. The
approximation scheme is based on a compromise between minimizing the
width of the shape function support and maximizing the information
entropy of the approximation. The \acrshort{lme} approximation
may be regarded as a regularization, or \textit{thermalization}, of
Delaunay triangulation which effectively resolves the degenerate cases
resulting from the lack of uniqueness or the triangulation. \acrshort{lme} basis functions possess many desirable properties for
meshfree algorithms. First of all, they are entirely defined by the
nodal set and the domain of analysis. They are also non-negative,
satisfy the partition of unity property, and provide an exact
approximation for related functions~\cite{Arroyo2006}. Important contributions on the Maximum-Entropy have been made by Sukumar and coworkers~\cite{Sukumar15} with Cell-based techniques and the ones carried within the \acrfull{otm} method. The latter methodology has been proven to have a good performance under
the dynamic regime by other researchers, being important the contributions of Li {\it et al.} (2012)~\cite{Li2012} and Navas {\it et al.}
(2018)~\cite{Navas:17b,Navas2018a} in the explicit regime and Navas
{\it et al.}~\cite{Navas2016,Navas2016b,Navas:17c} and Wriggers and
coworkers~\cite{Wriggers18} with implicit schemes. More
recently, under \acrshort{mpm} framework, the work made by Wobbes {\it et
  al.}(2020)~\cite{Wobbes2020}.\PNA{The proposed research delves into the benefits of
the regularization parameter, \gls{LME-beta}, and the analogy of the different obtained shape functions derive by the tuning of this parameter and the traditional \acrshort{mpm} ones.}

The aforementioned techniques are devoted to mitigate the
``grid crossing'' error. Nevertheless, in the presence of shock waves spurious,
numerical noises appear despite of the employment of these
techniques~\cite{Tran2019e}. These numerical inaccuracies, also known
as wiggles, arise due to inaccuracies in the time discretization technique.
A simple approach to face those spurious noises lies on the addition of nonphysical damping sources to the equilibrium equations. This
approach has been widely employed in this and many other numerical
techniques. To avoid introducing this nonphysical sources, many
researchers have proposed alternative time integration
schemes which reduce the presence of high frequency noises by
filtering them or increasing\PNA{somehow} the accuracy.\MMP{One of the first attempts was
the proposal of an implicit time integration scheme by Guilkey \&
Weiss \cite{Guilkey_2003}. More recently Wang {\it et al.}
\cite{Wang_2016}} \MP{mitigated} \MMP{these spurious noises by adding a non-viscous
damping to the linear momentum balance equation, and later Charlton
{\it et al.} \cite{Charlton_2017} extended this scheme to the
\acrshort{gimp} approach introducing the \acrfull{igimp}
method. However, the local damping introduced by 
\cite{Wang_2016} can totally over-damp the solution in time-dependent
simulations such as in consolidation process. In a recent publication,
Kularathna \& Soga \cite{Soga_2017} study an implicit treatment of the pressure in MPM algorithm to simulate material incompressibility
avoiding artificial pressure oscillations by applying
Chorin\textquotesingle s
projection. Within the explicit time integration schemes, Lu {\it et al.}\cite{LU_2018} introduced the time-discontinuous Galerkin method to control the spurious noises
propagation, and later Tran \& Solowski~\cite{Tran2019e}
proposed a generalised-$\alpha$ scheme for \acrshort{mpm} with
promising results but at the expense of increasing the computational
effort.} In this paper a less time consuming and high efficient explicit predictor-corrector integration method \MP{is}
proposed. It consists of an accommodation of the traditional \acrsull{npc} scheme, widely employed in Finite Element methods. This method
has been chosen among other suitable alternatives as those proposed
by Wilson {\it et al.} (1972)~\cite{Wilson1972} or Chung \& Hulbert
(1993)~\cite{Geranlized_alpha_1993} because its simplicity and good
performance dealing with solid dynamic problems under a meshfree
framework in~\cite{Navas2018a}.

@@

To accommodate this step to \acrshort{mpm} framework, it is necessary to get
the nodal values of the velocity and acceleration throughout a variational
recovery process where particles quantities are transferred to the
mesh nodes. This technique arises as a generalization of the super-convergent recovery
procedures described by Zienkiewicz \& Zhu \cite{ZZ1992_I} (\textit{ZZ})
in the context of \acrshort{fem}. In \acrshort{mpm}, Gauss quadratures are not employed. However, 
integrals are computed following the Riemann integral definition,
where each component of the summation corresponds to a particle of the
discretization. Also Bardenhagen \& Kober \cite{Bardenhagen2004}
proved that through this information-transference technique mass and momentum are conserved. So for a general particle variable $\Phi_p$, employing the
\textit{ZZ} technique, it is possible to get its nodal homologous $\Phi_I$  as:
\begin{equation}
  \label{eq:Variational-recovery}
   \Phi_I = \frac{m_p N_{Ip} \Phi_p}{m_I}.
 \end{equation}
 Therefore, to get an analogous expression for
 \eqref{eq:Predictor-velocity-I} in the context of \acrshort{mpm}, the
 procedure described in the equation \eqref{eq:Variational-recovery}
 is employed, obtainen the following expression:
 \begin{equation}
   \label{eq:Predictor-velocity-II}
   \vec{\tilde{v}}_I^{k+1} = \underbrace{\frac{N_{Ip}^{k} m_p
       \vec{v}_p^k}{m_I}}_{\vec{v}_I^{k}} + (1 - \gamma)\ \Delta t\  \underbrace{\frac{N_{Ip}^{k} m_p \vec{a}_p^k}{m_I}}_{\vec{a}_I^{k}}.
 \end{equation}
Nonetheless this way of computing the predictor stage can introduce
instabilities due to numerical cancellation likewise the original
Sulky algorithm. Thankfully, this can be avoided easily by the
equivalent formulation proposed as follows: 
\begin{equation}
  \label{eq:Predictor-velocity-II}
  \vec{\tilde{v}}_I^{k+1} = \frac{ N_{Ip}^{k} m_p (\vec{v}_p^k + (1 - \gamma)\ \Delta t\ \vec{a}_p^k)}{m_I}.
\end{equation}
This way of computing the nodal predictor is both numerically stable
and minimize the computational effort. Once nodal velocities are
obtained, the essential boundary conditions are imposed over \gls{dirichlet-boundary}. And in the
following, the ``classic'' \acrshort{mpm} algorithm continues to reach to the
equilibrium equation \eqref{eq:particle_balance_forces3}. Next, the
\textit{corrector} stage is introduced. \MP{As} nodal
velocities were obtained earlier, this step is computed in the same way as
in \acrshort{fem},
\begin{equation}
  \label{eq:Corrector-velocity}
  \vec{v}_{I}^{k+1} = \vec{v}_{I}^{pred} + \gamma\ \Delta t\ \frac{\vec{f}_{I}^{k+1}}{\tens{m}_I^{k+1}}.
\end{equation}
Finally updated particle \MP{accelerations, velocities and positions are updated as},
\begin{align}
  \label{eq:Update-lagrangian-pce}
        &\vec{a}_p^{k+1} = \frac{N_{Ip}^k\vec{f}_{I}^{k}}{\tens{m}_I^k}\\
      &\vec{v}_p^{k+1} = \vec{v}_p^n + \Delta t\
        \frac{N_{Ip}^k\
        \vec{f}_{I}^{k}}{\tens{m}_I^k}\\
      &\vec{x}_p^{k+1} = \vec{x}_p^n + \Delta t\
         N_{Ip}^k\ \vec{v}_{I}^{k} +
        \frac{1}{2}\Delta t^2\ \frac{N_{Ip}^k\
        \vec{f}_{I}^{k}}{\tens{m}_I^k}.
\end{align}
Notice that particle displacements are computed using the corrected
nodal velocities as well as the accelerations with the velocities
of the predictor. However, particle velocities and accelerations
are computed using the corrected velocities. Therefore here we share similarities
with the \textit{leapfrog \MP{scheme}} where position is not updated at
full time step, but the velocity is updated at half time steps. Notice
also that, with this approach, the calculation of nodal momentum values
are not required. Due to its simplicity, the proposed scheme can be implemented with minor modifications
over the standard forward Euler. The full implementation is summarized in the algorithm~\ref{algo:1}.

@@

\acrfull{lme} approximation schemes were
first introduced by Arroyo \& Ortiz (2006)\cite{Arroyo2006} and has been
recently tested under \acrshort{mpm} framework by Wobbes {\it et al.}
(2020)\cite{Wobbes2020}. \MP{The simulations presented in \cite{Wobbes2020} of \acrshort{mpm}
within \acrshort{lme} show considerably more accurate stress
approximations than traditional \acrshort{mpm} schemes.} \PNA{However, 
how the regularization parameter $\beta$ affects to the accuracy and
stability of the solution is not assessed deeply in that research, the tuning of this $\beta$ parameter allows to make the comparison of the accuracy against analogous traditional \acrshort{mpm} shape function.}

@@

application of this capability, where publication oscillations due to
excess of pore water pressure in consolidation problems are smoothed
out by using this technique. The employment of smoothing algorithm is
also straightforward in the fluid-solid interaction problems
\cite{Arduino_2018}.A proof of this statements is observed in
figure~\ref{fig:LME_MPM}, where the \acrshort{lme} adaptability is illustrate.}\\

@@

This section is devoted to test the ability of both predictor-corrector
time integration scheme and the Local \textit{Max-Ent} approximants to
overcome spurious oscillations due to the grid crossing and high
frequency loads under the context of \acrshort{mpm}. Two different test have
been adopted for this purpose: the benchmark proposed by Dyka \& Ingel
(1995)\cite{Dyka1995} and the test proposed in the PhD thesis of
Andersen (2009)\cite{thesis_Andersen_2009}. On the one hand, the accuracy of the \acrfull{npc}
scheme is compared to the standard \acrfull{fe} in order to assess the performance of the time integration schemes. On the other hand,  \acrshort{lme} solutions
are compared with those provided by \acrfull{ugimp} and Q4 shape
functions. To avoid some mesh-dependent issues, in all
calculations a regular background mesh was set. All simulations were
performed with in-house software.

@@

This benchmark was proposed\PNA{by Dyka~\cite{Dyka1995} since allows so study easily the capability}
of the proposed time integration algorithm to avoid velocity
field instabilities. It consists of a one-dimensional bar of a length
of 0.1333 meters, sketched in the figure~\ref{fig:Dyka_Bar}. The
boundary conditions are: displacements are constrained ($\vect{v}
\rvert_{x=L} = 0$) in the right border, \MP{being free on all other boundaries.} An initial velocity of $\vect{v}_o = - 5\ m/s$ is given to the
left quarter of the bar. Finally, the elastic parameters \MP{chosen} for this test are:
\begin{itemize} 
\item  Density : $7833\ kg/m3$
\item  Poisson ratio : $0$
\item  Elastic modulus : $200 \cdot 10^9\ Pa$
\end{itemize}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}\sidecaption
  \centering
  \resizebox{\hsize}{!}{
    \input{Figures/Dyka_bar.tex}}
  \caption{Geometrical description of the Dyka \cite{Dyka1995} bar.}
  \label{fig:Dyka_Bar}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\MP{In the proposed exmple, the simulation extends from time zero up to 0.00001 s.
This time interval allows the elastic wave to travel a distance of 2.6
lengths of the bar}. For the spatial
discretization, a set of seven nodal mesh sizes (0.1, 0.3325, 0.5,
1.0, 3.3325, 6.665, 10.0 millimeters) are considered. For each element a number of
four particles was selected. In the initial layout, particles are
\MP{located} the exact quadrature points of a linear quadrilateral, with
the exception of the \acrshort{ugimp} simulation, where gaps or overlap between
voxels of each particle are not allowed. In those cases, each particle
occupies the center of each cell quarter. For all simulations, time step
is controlled by a Courant-Friedrichs-Levy condition of 0.1, were the adopted
celerity is computed as:
\begin{equation}
  \label{eq:Cel}
  Cel = \max\{\max_{p \in \Omega_p}\{ \vect{v}_p \} , \max_{p \in \Omega_p}\{ \sqrt{\frac{E_p}{\rho_p}} \} \}.
\end{equation}
An important consideration regarding modellization concerns the
background mesh. Notice that free border of the bar has a maximum
horizontal displacement of 0.03 millimeters, therefore 
a computational domain with an extra gap of 0.03 millimeters is
required in order to accommodate the unconstrained displacement of the
particles in the left border of the bar. Naturally this problem arises
when the mesh size is small enough that relative displacement of the
particles is larger then the distance to the border, so grid crossing
phenomena could appear even in those cases with infinitesimal
displacements.

@@

A first \MP{comparison} between the \MP{obtained in} both time integration
schemes is plotted in figure \ref{fig:Dyka-NPC-FE}. It demonstrates the
superior performance of the \acrshort{npc} \textit{versus} the \acrshort{fe}. In the \acrshort{npc} the
spurious oscillations are quickly mitigated in the first time steps,
and the error is not propagated in time, opposite to the results of the \acrshort{fe} scheme, where the
simulation becomes unstable after $6E^{5}$ seconds.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}\sidecaption
  \centering
  \resizebox{0.9\hsize}{!}{
    \includegraphics[width=\textwidth]{./Figures/Velocity_FE_vs_PCE_CFL_01}
  }
  \caption{\MP{comparison of \acrshort{npc} and \acrshort{fe}
      performances}: In the picture the velocity evolution at the point in the bar left side
    is plotted.}
  \label{fig:Dyka-NPC-FE}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
